---
title: "R Notebook"
output: html_notebook
---

```{r}

library(openxlsx)
library(tidyverse)
library(gt)
library(dplyr)
library(tidyr)
data <- read.xlsx("Kopie von Patient_characteristics3.xlsx", 
                  sheet = "Adverse Events") %>%
  mutate(Relatedness = str_remove(Relatedness, " $")) # remove trailing spaces

```


All patients with any AE (all grades):

```{r}

data %>%
  select(patient) %>%
  unique() %>%
  nrow()

```

All patients with any AE (grade 3 to 5):

```{r}

data %>%
  filter(Grade %in% c("G3", "G4", "G5")) %>%
  select(patient) %>%
  unique() %>%
  nrow()

```

Treatment related (all grades)

```{r}

data %>%
  filter(Relatedness %in% c("related", "possibly related")) %>%
  select(patient) %>%
  unique() %>%
  nrow()

```

Treatment related (grades 3 to 5)

```{r}

data %>%
  filter(Relatedness %in% c("related", "possibly related")) %>%
  filter(Grade %in% c("G3", "G4", "G5")) %>%
  select(patient) %>%
  unique() %>%
  nrow()

```

Frequency of individual AEs:
```{r}

# any grade
ae_freq_all <- data %>% 
  select(patient, Adverse.Events) %>%
  unique() %>%
  group_by(Adverse.Events) %>%
  summarise(all_grades = length(unique(patient))) %>% 
  arrange(desc(all_grades))

# grades 1 to 2
ae_freq_mild <- data %>%
  filter(Grade %in% c("G1", "G2")) %>%
  select(patient, Adverse.Events) %>%
  unique() %>%
  group_by(Adverse.Events) %>%
  summarise(g12 = length(unique(patient))) %>% 
  arrange(desc(g12))

# grades 3 to 5
ae_freq_severe <- data %>%
  filter(Grade %in% c("G3", "G4", "G5")) %>%
  select(patient, Adverse.Events) %>%
  unique() %>%
  group_by(Adverse.Events) %>%
  summarise(g35 = length(unique(patient))) %>% 
  arrange(desc(g35))

# Combine above in one data.frame
ae_freqs_together <- merge(ae_freq_all, ae_freq_mild, by = 1, all = TRUE) %>% merge(., ae_freq_severe, by = 1, all = T) %>% arrange(desc(all_grades))

# Save as xlsx
write.xlsx(ae_freqs_together, "AE_frequencies.xlsx")

```

```{r}
# Define total patient set from your data if needed
total_patients <- length(unique(data$patient))  # optional

# Base data: add grade_group
data <- data %>%
  mutate(
    grade_group = case_when(
      Grade %in% c("G1", "G2") ~ "g12",
      Grade %in% c("G3", "G4", "G5") ~ "g35",
      TRUE ~ NA_character_
    )
  )

# Create three labeled subsets
# Create grouped datasets
all_ae <- data %>%
  filter(!is.na(grade_group)) %>%
  mutate(group = "All Adverse Events")

related_ae <- data %>%
  filter(!is.na(grade_group), Relatedness == "related") %>%
  mutate(group = "Treatment-related")

#possibly_related_ae <- data %>%
#  filter(!is.na(grade_group), Relatedness %in% c("related", "possibly related")) %>%
#  mutate(group = "Treatment-related or possibly related")

# Combine all rows
summary_data <- bind_rows(all_ae, 
                          related_ae, 
                          #possibly_related_ae
                          )

# g12 and g35: per patient, per group, per grade_group
g12_g35_counts <- summary_data %>%
  distinct(patient, group, grade_group) %>%
  count(group, grade_group, name = "n") %>%
  pivot_wider(names_from = grade_group, values_from = n, values_fill = 0)

# all_grades: per patient per group (regardless of grade group)
all_grades_counts <- summary_data %>%
  distinct(patient, group) %>%
  count(group, name = "all_grades")

# Join them cleanly
summary_rows <- full_join(all_grades_counts, g12_g35_counts, by = "group") %>%
  replace_na(list(g12 = 0, g35 = 0)) %>%
  rename(Adverse.Events = group) %>%
  select(Adverse.Events, all_grades, g12, g35)
```

```{r}
# Copy the original dataframe
ae_table <- rbind(summary_rows, ae_freqs_together)

# Helper function to format numbers with percentage, or "-" for NAs
format_with_percent <- function(x) {
  if (is.na(x)) {
    return("-")
  } else {
    return(sprintf("%d (%.0f%%)", x, 100 * x / total_patients))
  }
}

# Apply formatting column-wise
ae_table$all_grades <- sapply(ae_table$all_grades, format_with_percent)
ae_table$g12        <- sapply(ae_table$g12,        format_with_percent)
ae_table$g35        <- sapply(ae_table$g35,        format_with_percent)

# Create the gt table
gt_table <- ae_table %>%
  gt() |>
  opt_stylize(style = 1, , color = "gray") %>%
  cols_label(
    Adverse.Events = "Adverse Events",
    all_grades     = "All Grades",
    g12            = "Grade 1-2",
    g35            = "Grade 3-5"
  ) %>%
  cols_align(
    align = "center",
    columns = c(all_grades, g12, g35)  # target all but first column
  )

# Show the table
gt_table
```

```{r}
```


```{r}
```


```{r}
gtsave(gt_table, "ae_table_13-08-26.pdf")
```


```{r}
```

Cemi realted, all grades

```{r}

data %>%
  filter(Medication_related == "Cemiplimab") %>%
  filter(Grade %in% c("G3", "G4", "G5")) %>%
  select(patient) %>%
  unique() %>%
  nrow()

```

Cemi realted, grades 3 to 5

```{r}

data %>%
  filter(Medication_related == "Cemiplimab") %>%
  filter(Grade %in% c("G3", "G4", "G5")) %>%
  select(patient) %>%
  unique() %>%
  nrow()

```

Soni related, all grades 

```{r}

data %>%
  filter(Medication_related == "Sonidegib") %>%
#  filter(X3 %in% c("G3", "G4", "G5")) %>%
  select(patient) %>%
  unique() %>%
  nrow()

```

Soni related, grades 3 to 5

```{r}

data %>%
  filter(Medication_related == "Sonidegib") %>%
  filter(Grade %in% c("G3", "G4", "G5")) %>%
  select(patient) %>%
  unique() %>%
  nrow()

```

Related to both, all grades
(I understand this to mean that a patient has an AE related to Cemi, and 
also an AE related to Soni. They don't have to be the same AE.)

```{r}

data %>%
  filter(!is.na(Medication_related)) %>%
  select(patient, Medication_related) %>%
  unique() %>%
  group_by(patient) %>%
  summarise(num_drugs = length(unique(Medication_related))) %>%
  filter(num_drugs == 2) %>%
  nrow()


```


Related to both, grades 3 to 5

```{r}

data %>%
  filter(!is.na(Medication_related)) %>%
  filter(Grade %in% c("G3", "G4", "G5")) %>%
  select(patient, Medication_related) %>%
  unique() %>%
  group_by(patient) %>%
  summarise(num_drugs = length(unique(Medication_related))) %>%
  filter(num_drugs == 2) %>%
  nrow()


```

Unknown which drug, all grades 
(Here is "related" or "possibly related" but NA in Medication_related)

```{r}

data %>%
  filter(is.na(Medication_related)) %>%
  filter(Relatedness %in% c("related", "possibly related")) %>%
#  filter(X3 %in% c("G3", "G4", "G5"))
  select(patient) %>%
  unique() %>%
  nrow()


```

Unknown which drug, grades 3 to 5
(Here is "related" or "possibly related" but NA in Medication_related)

```{r}

data %>%
  filter(is.na(Medication_related)) %>%
  filter(Relatedness %in% c("related", "possibly related")) %>%
  filter(X3 %in% c("G3", "G4", "G5")) %>%
  select(patient) %>%
  unique() %>%
  nrow()


```
Response summary table
```{r}
summarydata <- read.delim("summary.tsv",
                     header = TRUE, sep = "\t")

```

```{r}
sum_table <- summarydata %>%
  gt() |>
  opt_stylize(style = 1, , color = "gray") %>%
  cols_label(
    Response.type = "Response Type",
    Number.of.Patients = "Number of Patients",
    Proportion = "Proportion (%)") %>%
    tab_header(
    title = md("Best response any time between treatment start and week 26"),
    subtitle = md("Patients (n=20)"))

sum_table
```

```{r}
gtsave(sum_table, "summary_table.pdf")
```

Supp Table 1
```{r}
supp1 <- read.delim("SuppTable/SuppTable1.tsv",
                     header = TRUE, sep = "\t")
colnames(supp1)
```

```{r}
supp1_table <- supp1 %>%
  gt() |>
  opt_stylize(style = 1, , color = "gray") %>%
  cols_label(
    Patient.ID = "ID",
    Sex = "Sex",
    Age = "Age (years)",
    Comorbidities = "Hematological malignancies and genetic diseases",
    Tumor.Stage = "Tumor Stage",
    prior.local.therapy = "Prior Local Therapy",
    prior.systemic.therapy = "Prior Systemic Therapy"
    ) |>
  tab_options(
    page.orientation = "landscape",
    page.width = "11.7in",
    page.height = "8.3in",
    table.font.size = px(10)
  )|> tab_header(
    title = md("Supplementary Table 1"),
    subtitle = md("Patient characteristics, including hematological malignancies and genetic diseases, local and systemic treatments.")
  ) |> opt_align_table_header(align = "left")

supp1_table
```

```{r}
gtsave(supp1_table, "supp1_table.pdf")
```
Supp Table 2
```{r}
supp2 <- read.delim("SuppTable/SuppTable2.tsv",
                     header = TRUE, sep = "\t")
colnames(supp2)
```

```{r}
supp2_table <- supp2 %>%
  gt() |>
  opt_stylize(style = 1, , color = "gray") %>%
  cols_label(
    Patient.ID = "ID",
    Prior.HHI. = "Prior HHI",
    Best.response.to.prior.HHI = "Best response to prior HHI",
    Reason.for.discontinuation.of.prior.HHI = "Reason for discontinuation of prior HHI",
    Washout.duration.before.study.entry = "Washout duration before study entry (weeks)",
    Duration.of.Vismodegib = "Duration of vismodegib (weeks)",
    Duration.of.Sonideib. = "Duration of Sonidegib (weeks)"
    ) |>
  tab_options(
    page.orientation = "landscape",
    page.width = "11.7in",
    page.height = "8.3in",
    table.font.size = px(10)
  )|> tab_header(
    title = md("Supplementary Table 2"),
    subtitle = md("Data on prior HHI treatment.")
  ) |> opt_align_table_header(align = "left")

supp2_table
```

```{r}
gtsave(supp2_table, "supp2_table.pdf")
```


```{r}
supp3 <- read.delim("SuppTable/SuppTable3.tsv",
                     header = TRUE, sep = "\t")
colnames(supp3)
```

```{r}
supp3_table <- supp3 %>%
  gt() |>
  opt_stylize(style = 1, , color = "gray") %>%
  cols_label(
    ID = "ID",
    Response.at.week.26 = "Response at week.26",
    Best.response.anytime.between.treatment.start.and.week.26 = "Best response anytime between treatment start and week 26"
) |>
  tab_options(
    page.orientation = "landscape",
    page.width = "11.7in",
    page.height = "8.3in",
    table.font.size = px(10)
  )|> tab_header(
    title = md("Supplementary Table 3"),
    subtitle = md("Responses at week 26 and best response by week 26.")
  ) |> opt_align_table_header(align = "left")

supp3_table
```

```{r}
gtsave(supp3_table, "supp3_table.pdf")
```

```{r}
supp4 <- read.delim("SuppTable/SuppTable4.tsv",
                     header = TRUE, sep = "\t")
colnames(supp4)
```

```{r}
supp4_table <- supp4 %>%
  gt() |>
  opt_stylize(style = 1, , color = "gray") %>%
  cols_label(
    Pat.ID = "ID",
    Cemiplimab.cycles.received = "Cemiplimab cycles received",
    Sonidegib.cycles.received = "Sonidegib cycles received",
    Completed.26.Wk.regimen. = "Completed 26wk regimen",
    Dose.interruption = "Dose interruption (Drug/reason)",
    Reason.for.early.study.discontinuation = "Reason for early study discontinuation",
    Dose.Reduction = "Dose Reduction",
    Post.26.week.Status...Treatment = "Post 26wk Status / Treatment"
    ) |>
  tab_options(
    page.orientation = "landscape",
    page.width = "11.7in",
    page.height = "8.3in",
    table.font.size = px(10)
  )|> tab_header(
    title = md("Supplementary Table 4"),
    subtitle = md("Dose intensity, interruptions, reductions and post study status")
  ) |> opt_align_table_header(align = "left")

supp4_table
```

```{r}
gtsave(supp4_table, "supp4_table.pdf")
```

